
% --- This goes into sections/00_preliminaries.tex ---

\chapter*{Acknowledgements}

At the conclusion of this enriching internship experience, I wish to express my sincere and profound gratitude to all those who contributed, directly or indirectly, to the successful completion of this project and the writing of this report.

My deepest gratitude goes first and foremost to my supervisor at the Office National Des Aéroports (ONDA), \textbf{Mr. Darnag RACHIDE}. His constant availability, invaluable technical guidance, and insightful feedback were instrumental from the project's inception to its final deployment. I am particularly grateful for the trust he placed in me and for the autonomy he granted, which allowed me to explore and implement advanced machine learning solutions. His expertise was a guiding light in navigating the complexities of the flight dataset and in refining the architecture of the predictive models.

I extend my sincere thanks to the entire management of \textbf{ONDA} for providing me with this exceptional opportunity to work on a project of such practical and strategic importance. The professional and stimulating environment was highly conducive to learning and professional growth.

I would also like to thank my colleagues within the department for their warm welcome and collaborative spirit. Their support, the insightful technical discussions, and the positive atmosphere they fostered made this internship a truly valuable human and professional experience.

My gratitude also extends to the entire faculty of the \textbf{National School of Computer Science and Systems Analysis (ENSIAS)}, and particularly to the teaching staff of the \textbf{Artificial Intelligence (2IA)} program. The solid theoretical foundation in machine learning, data processing, and software engineering acquired during my studies was the bedrock upon which this project was built.

Finally, I wish to express my heartfelt thanks to my family and friends for their unwavering support, encouragement, and patience throughout this journey.
\chapter*{Abstract}
This report details the work accomplished during my internship at the Office National Des Aéroports (ONDA). The primary objective of this project was to develop an end-to-end Flight Traffic Analysis and Prediction Platform. This involved processing extensive historical flight data, engineering relevant features to capture seasonality and the impact of the COVID-19 pandemic, and training multiple machine learning models. The final deliverable is a fully interactive web dashboard, built with Dash Plotly, that allows non-technical users to explore historical trends and generate on-demand forecasts for future flight volumes.
\vspace{1cm}

\noindent\textbf{Keywords:} Machine Learning, Time Series Forecasting, Data Visualization, Python, Scikit-learn, LightGBM, Dash Plotly, Air Traffic Management.python

\chapter*{Résumé}

\setlength{\parindent}{1em} % normal paragraph indentation

\noindent
Ce rapport détaille les travaux réalisés lors de mon stage à l'Office National Des Aéroports (ONDA). 
L'objectif principal de ce projet était de développer une plateforme complète d'analyse et de prédiction du trafic aérien.

Cela a impliqué le traitement de données de vol historiques, la création de caractéristiques pertinentes pour capturer la saisonnalité et l'impact de la pandémie de COVID-19, et l'entraînement de plusieurs modèles de machine learning.

Le livrable final est un tableau de bord web interactif, construit avec Dash Plotly, qui permet aux utilisateurs non techniques d'explorer les tendances historiques et de générer des prévisions à la demande pour les volumes de vols futurs.

\vspace{1cm}

\noindent\textbf{Mots-clés :} 
Apprentissage automatique, Prévision de séries temporelles, Visualisation de données, Python, Scikit-learn, LightGBM, Dash Plotly, Gestion du trafic aérien.


\tableofcontents
\listoffigures
\listoftables
\cleardoublepage
